### 2016-06-18 R/ITHIM.Rmd
Update to ITHIM.Rmd

## R Package _ITHIM_
This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
library("ggplot2")
library("reshape2")
library("plyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
library("devtools")
install("~/ITHIM/")
# install_github("syounkin/ITHIM", force = TRUE)
library("ITHIM")
date()
```

### Running in R

```{r instantiate}
ITHIM.baseline <- list(
  parameters = parameters <- createParameterList(vision = "baseline", region = "national"),
  means = means <- computeMeanMatrices(parameters),
  quintiles = quintiles <- getQuintiles(means, region = "national")
)
ITHIM.scenario <- list(
  parameters = parameters <- createParameterList(vision = "scenario", region = "national"),
  means = means <- computeMeanMatrices(parameters),
  quintiles = quintiles <- getQuintiles(means, region = "national")
)
```
Now we can compute the attributable fraction (AF).

```{r RRs}
comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

```{r show}
with(comparitiveRisk, 100*round(AF$BreastCancer,4))
```

@@

```{r parlist}
AF.sum <- numeric(1e3); i <- 0;
muwt.vec <- seq(107.1,150,length.out=length(AF.pseudo))
#muct.vec <- seq(2.7,3.0,length.out=length(AF.pseudo))
for(muwt in muwt.vec){
i <- i+1
parameters <- createParameterList(baseline = FALSE)

parameters <- setParameter(parName="muwt", parValue = muwt, parList = parameters)
ITHIM.scenario <- list(
  parameters = parameters <- parameters,
  means = means <- computeMeanMatrices(parameters),
  quintiles = quintiles <- getQuintiles(means)
)
comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
AF.sum[i] <- with(comparitiveRisk, sum(100*AF$BreastCancer))
}
D <- data.frame(mu.wt = muwt.vec,AF.sum=AF.sum)
p <- ggplot(D, aes(x = mu.wt, y = AF.sum))
p + geom_point()
```

```{r AF, echo = FALSE}
with(comparitiveRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
```
```{r ap}
k <- createAirPollutionRRs()
```

### Relative Risk Model
```{r explore}
names(ITHIM.baseline$parameters)
```
```{r RRFigure, echo = FALSE}
phi <- function(x,k=0.5,phi0=comparitiveRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
```

@@

```{r plot2, echo = FALSE, fig.width = 12}
plotRR(comparitiveRisk$RR.baseline$Depression,comparitiveRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
```

```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
plotRR(comparitiveRisk$RR.baseline$BreastCancer,comparitiveRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
```

```{r plot4, echo = FALSE, fig.height = 12}
D<-melt(comparitiveRisk$AF)
names(D)<- c("age","sex","AF","disease")
p <- ggplot(D, aes(age,AF)) + geom_bar(aes(fill=sex), stat = "identity") + facet_grid( disease ~ .) + coord_cartesian(ylim = c(0, 0.1))
p
```
### Road Injury Array
```{r road}
baselineInjuries <- array(dim=c(6,7,3,2))
injury <- c("fatal","serious")
road <- c("minor road","major road","motorway")
vehicle <- c("walk","cycle","bus","car","HGV","mbike")
dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)

@@

```
# Code Graveyard

@@

```{r saveMortality, echo = FALSE, eval = FALSE}
saveRDS(object = Data.Mortality, file = "./data/Data.Mortality.rds")
saveRDS(object = Data.Mortality.Table, file = "./data/Data.Mortality.Table.rds")
```

### Appendix

>>>>

## R Package _ITHIM_
<!-- This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.-->

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
library("ggplot2")
library("reshape2")
library("plyr")
#install("~/ITHIM/")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
library("devtools")
install_github("syounkin/ITHIM")
library("ITHIM")
```

### Bay Area

First we define the baseline and scenario objects.

```{r instantiate}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
```

Next we compare the scenario to the baseline using compareModels().

```{r RRs}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

#### Deaths
```{r show, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### DALY
```{r show2, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLD
```{r show3, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show4, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

### National

```{r instantiate2}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
```

```{r RRs2}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

#### Deaths
```{r show222, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### DALY
```{r show22, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLD
```{r show23, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show24, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

### Creating Different Visions

```{r vision}
ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
```
#### YLL
```{r show233, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

### Interrogate Parameters

```{r params}
names(ITHIM.baseline$parameters)
ITHIM.baseline$parameters$NonTravelMETs
ITHIM.baseline$parameters$Rwt
```

<!--

# Code Graveyard

### Relative Risk Model
```{r explore}
names(ITHIM.baseline$parameters)
```
```{r RRFigure, echo = FALSE}
phi <- function(x,k=0.5,phi0=comparativeRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
```

### Tinkering

@@

```{r parlist}
AF.sum <- numeric(1e3); i <- 0;
muwt.vec <- seq(107.1,150,length.out=length(AF.sum))
#muct.vec <- seq(2.7,3.0,length.out=length(AF.sum))
for(muwt in muwt.vec){
i <- i+1
parameters <- createParameterList(baseline = FALSE)

parameters <- setParameter(parName="muwt", parValue = muwt, parList = parameters)
ITHIM.scenario <- list(
  parameters = parameters <- parameters,
  means = means <- computeMeanMatrices(parameters),
  quintiles = quintiles <- getQuintiles(means)
)
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
AF.sum[i] <- with(comparativeRisk, sum(100*AF$BreastCancer))
}
D <- data.frame(mu.wt = muwt.vec,AF.sum=AF.sum)
p <- ggplot(D, aes(x = mu.wt, y = AF.sum))
p + geom_point()
```

```{r AF, echo = FALSE}
with(comparativeRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
```

@@

```{r plot2, echo = FALSE, fig.width = 12}
plotRR(comparativeRisk$RR.baseline$Depression,comparativeRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
```

```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
plotRR(comparativeRisk$RR.baseline$BreastCancer,comparativeRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
```

```{r plot4, echo = FALSE, fig.height = 12}
D<-melt(comparativeRisk$AF)
names(D)<- c("age","sex","AF","disease")
p <- ggplot(D, aes(age,AF)) + geom_bar(aes(fill=sex), stat = "identity") + facet_grid( disease ~ .) + coord_cartesian(ylim = c(0, 0.1))
p
```
### Road Injury Array
```{r road}
baselineInjuries <- array(dim=c(6,7,3,2))
injury <- c("fatal","serious")
road <- c("minor road","major road","motorway")
vehicle <- c("walk","cycle","bus","car","HGV","mbike")
dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)

@@

@@

```{r saveMortality, echo = FALSE, eval = FALSE}
saveRDS(object = Data.Mortality, file = "./data/Data.Mortality.rds")
saveRDS(object = Data.Mortality.Table, file = "./data/Data.Mortality.Table.rds")
```
-->
### Appendix
